---
title: "Data Exploration Project"
author: "Bethany Locke"
date: "2/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

In this code chunk, load in all of the libraries that will be needed:

```{r cars}
library(tidyverse)
library(dplyr)
library(vtable)
library(jtools)
library(fixest)
library(car)
library(wooldridge)
library(purrr)
library(lubridate)

```
## Research Question

The College Scorecard was released in September 2015. Among colleges that predominantly grant bachelor's degrees, our research question is exploring if the release of the Scorecard shifts student interest to high-earnings colleges relative to low-earnings ones (as proxied by Google searches for keywords associated with those colleges).

## Data Wrangling is in Cleaning Script file

```{r pressure, include=FALSE}
# Identifying all CSV files in folder, storing all files in list, combining data sets into one data set

trends <- list.files(path = "Raw Data", 
                       pattern = "trends_up_to_", full.names = TRUE) %>%
  lapply(read_csv) %>%
  bind_rows

# Reading in the Scorecard and IDName CSV files

Scorecard <- read.csv(file = 'Most+Recent+Cohorts+(Scorecard+Elements).csv')
IDName <- read.csv(file = 'id_name_link.csv')

# Removing duplicate university and/or college names (schname)

IDName <- IDName %>%
  group_by(schname) %>%
  mutate(N = n()) %>%
  filter(N == 1)

# Converting column names in Scorecard to match lowercase 
names(Scorecard) <- tolower(names(Scorecard))

# Combining IDName and trends into IDTRends by the school's name

IDTrends <- merge(x = IDName, y = trends, by = "schname")

# Filtering the Scorecard to display colleges/universities that predominantly grant Bachelor degrees as BachelorData 

filter(Scorecard, preddeg == 3)
Scorecard <- Scorecard %>% select(unitid, opeid, preddeg, md_earn_wne_p10.reported.earnings) %>%
  filter(preddeg == 3)

# Merging the previous EarnngsColleges scorecard with IDTrends as IDEarningsColleges before creating the binary

IDEarningsColleges <- merge(x = IDTrends, y = Scorecard, by = 'opeid')

# Creating the binary for high-earnings and low-earnings colleges/universities that predominantly grant Bachelor degrees. $60,000 is the identified median reported earnings that would set the binary.

IDEarningsColleges$earningshigh <- ifelse(IDEarningsColleges$md_earn_wne_p10.reported.earnings >= 60000, "1", "0")

# Standardizing IDEarningsColleges as TrendsBachelorMonthGrouped by month, opeid, and index and displaying the dates correctly using lubridate.

TrendsBachelorMonthGrouped <- IDEarningsColleges %>%
  mutate(date = as.Date(str_sub(monthorweek, 1, 10))) %>%
  group_by(schname, keyword) %>%
  mutate(index_std = (index - mean(index,na.rm = TRUE))/sd(index, na.rm = TRUE)) %>%
  group_by(month=floor_date(date, "month"), opeid) %>%
  summarize(index = mean(index, na.rm = TRUE))

# Joining BachelorData and IDName

BachelorIDData <- left_join(IDName, Scorecard, by = 'opeid')

# Officially cleaned data before dropping NA values

CleanOfficial <- left_join(TrendsBachelorMonthGrouped, BachelorIDData, by = 'opeid')

# Only looking at distinct columns drom IDEarningsColleges
IDEarningsSort <- IDEarningsColleges %>% select(earningshigh, opeid)
DistinctIDEarningsSort <- distinct(IDEarningsSort, opeid, earningshigh)

# Dropping all NA values
CleanReal <- left_join(CleanOfficial, DistinctIDEarningsSort)
CleanReal <- CleanReal %>% drop_na()

```


## Data Analysis and Regression

```{r}

# Regression model designed to answer the research question using difference in difference (DID) analysis

df <- CleanReal %>%
  mutate(after = month >= as.Date('2015-09-01'),
         treated = earningshigh > 0)

newdf <- df %>% 
  group_by(after, treated) %>%
  summarise(proportion_treated = mean(index),.groups = 'drop')

means <- df %>% 
  group_by(after, treated) %>%
  summarise(proportion_treated = mean(index),.groups = 'drop') %>%
  pull(proportion_treated)

# The difference between index for high-earnings colleges is -5.389817
(means[4] - means[2])

# The difference between index for low-earnings colleges is -4.26886
(means[3] - means[1])

# The difference in means is -1.120957, meaning that the difference in means for low-earnings college is less than the difference in means for high-earnings colleges.

(means[4] - means[2]) - (means[3] - means[1])

# Regression

lm(index ~ after*treated, data = df) %>%
  export_summs(digits = 3)

# Creating a graph displaying the high- and low-earnings colleges index over time

ggplot(df, aes(month, index, color = earningshigh, group = earningshigh)) +
    stat_summary(geom = 'line') +
    geom_vline(xintercept = as.Date('2015-09-01')) + theme_minimal()

```

## Interpretation and Write-Up

# Why you are performing the analysis you are performing, and the choices you made in putting it together?

In this Data Exploration project, we are performing a difference in difference analysis referring to the College/University site visitation both before and after the College Scorecard’s release in September 2015. We are using a difference in difference (DID) analysis because it addresses using within variation in order to identify the effect we want (release of the scorecard). 

One of the choices made was for identifying high-earning and low-earning colleges that predominately grant bachelor’s degrees. When creating this binary, we needed to indicate median reported earnings that would differentiate between high-earning and low-earning colleges. Although earning a bachelor’s degree in a certain study may pay more than other studies, a calculated average for recent years is about $60,000 (ZipRecruiter, CNBC, Statista). 

# Explain how your analysis addresses the research question.

The research question we are addressing is for the colleges that predominantly grant bachelor’s degrees, and if the release of the Scorecard shifts student interest to high-earnings colleges relative to low-earning ones. This metric can be evaluated by the proxied Google searches for keywords associated with those colleges. The difference in difference analysis addresses the research question because it displays the difference in means of high-earning colleges both before and after the scorecard release, and low-earning colleges both before and after the release. Because of parallel trends holding, these two means can be compared.

# Any additional analyses you did that led you to design your main analysis that way?

Some additional analyses that altered the design of the main analysis was how the graph was displayed. Graphing every tick mark for month is too busy, so year was graphed instead with the inclusion of an x-intercept September 2015 to denote when the College Scorecard was released. 

# Explain what we should conclude, in real world terms, based on the results

Based on our conclusions, the release of the College Scorecard did not shift student interest from high-earnings colleges to low-earnings ones. The difference in mean viewership of low-earnings colleges is less than that of high-earnings colleges (-4.27 in comparison to -5.39). In actuality, the College Scorecard has altered student interest from less overall college/university site visitation (most prevalent with high earning colleges than low earning colleges). Even though low earning colleges had less site visitation after the scorecard release, the mean difference is less distinguishable than high earning colleges. The graph portrays viewership of low-earnings colleges coming up to par with high-earnings colleges pending after the scorecard's release in September 2015. This can be described by the difference between the viewership levels between high- and low-earnings colleges becoming closer together post September 2015.  This results can conclude that the College Scorecard presents more holistic information of a range of colleges for people to view at once following the release in September. The College Scorecard may aim to reduce unequal viewership of college/university sites in relation to earnings with a more holistic summary of their scorecard. This gives a chance for low-earning colleges to gain more attention.

After the College Scorecard's release, a high-earnings college is 0.866 percentage points more likely to be viewed than low-earnings colleges. Regardless of if the college is high- or low-earnings and before the College Scorecard release, the average index viewership for a given Bachelor-granting college is 48.047. After the College Scorecard release for low-earnings colleges, the difference of means is -4.269, which is a smaller mean difference than that of high-earnings colleges (-5.389). The difference in difference analysis depicts that the difference in means is -1.121, which shows that low-earnings colleges has a greater difference in means than high-earnings colleges.

The findings from this analysis indicate that colleges should not expect a drastic increase in proxied searches after the College Scorecard's release. Rather, college/universities can invest their budget in marketing efforts that will drive site traffic, since the College Scorecard release is not indicative of increased traffic.